---
title: "Project 2"
author: "Ryan Thomas"
date: "2024-11-15"
output: html_document
---

## R Markdown

## **1. Introduction**

The topic of my project is a dataset about the seven major social media platforms such as Instagram and Snapchat, and how the different types of phone have different price points and market value. I chose this data set because it something that I have a lot of interest in and I have had most of these phone models. I have a deep understanding of the data and can make statistics on relationships between variables easier. The link to the data set is provided below: [https://www.kaggle.com/datasets/willianoliveiragibin/phone-information-2024]()

## **2. Data Preparation**

Uploading the data was an easy process. I had found the dataset on Kaggle and then I did the process of going to the source which had the table listed and then I downloaded the file in the form of a csv. The file was downloaded successfully and it can be replicated

## **3. Variable Descriptions**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
phone <- read.csv("/Users/ryanthomas/downloads/DS/phone_info.csv")
```

```{r}
library(knitr)

# Load the dplyr package
library(dplyr)

# Select relevant columns from the dataset
cleaned_data <- phone %>%
  select(phone_brand, phone_model, price_USD, storage, ram, Display_Size, OS)


# Variable description table
variable_descriptions <- data.frame(
  Variable = c("phone_brand", "phone_model", "price_USD", "storage", "ram", "Display_Size", "OS"),
  Type = c("Character", "Character", "Numeric", "Numeric", "Numeric", "Numeric", "Character"),
  Description = c(
    "Brand of the smartphone",
    "Specific model of the smartphone",
    "Price of the phone in USD",
    "Storage capacity in GB",
    "RAM in GB",
    "Screen size in inches",
    "Operating System (e.g., iOS, Android)"
  )
)

# Display the variable descriptions table
kable(variable_descriptions, col.names = c("Variable", "Type", "Description"), format = "markdown")

# Create new variables
phone$price_per_gb <- phone$price_USD / phone$storage
phone$performance_score <- phone$ram * phone$storage

# View the first few rows of the updated dataset
head(phone)
```
## **4. Univariate Analyse**
```{r}
# Load necessary library
library(ggplot2)
#these codes are all inline codes that are separated by the + symbol in different blocks that are continuous
# Create a histogram for price_USD
ggplot(phone, aes(x = price_USD)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_density(aes(y = ..count..), color = "blue", size = 1) +  
  labs(
    title = "Distribution of Prices (USD)",
    x = "Price (USD)",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  # this code helps make the plot title move to the middle
    axis.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed")
  )
# Histogram for RAM
ggplot(phone, aes(x = ram)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of RAM", x = "RAM (GB)", y = "Frequency") +
  theme_minimal()

# Histogram for Battery Capacity
ggplot(phone, aes(x = BATTERY)) +
  geom_histogram(binwidth = 500, fill = "green", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Battery Capacity", x = "Battery Capacity", y = "Frequency") +
  theme_minimal()

# Histogram for Display Size
ggplot(phone, aes(x = Display_Size)) +
  geom_histogram(binwidth = 0.5, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Display Size", x = "Display Size (inches)", y = "Frequency") +
  theme_minimal()
```
## **5. Multivariate Analyse**
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# 1. Scatter Plot: Display_Size vs. BATTERY with linear regression
ggplot(phone, aes(x = Display_Size, y = BATTERY)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Display Size vs. Battery Capacity",
       x = "Display Size (inches)",
       y = "Battery Capacity (mAh)") +
  theme_minimal()

# Calculate correlation
correlation <- cor(phone$Display_Size, phone$BATTERY, use = "complete.obs")
print(paste("Correlation between Display Size and Battery Capacity:", round(correlation, 2)))

# 2. Box Plot: RAM across Price Ranges
ggplot(phone, aes(x = price_range, y = ram, fill = price_range)) +
  geom_boxplot() +
  labs(title = "RAM Distribution Across Price Ranges",
       x = "Price Range",
       y = "RAM (GB)") +
  theme_minimal()

# 3. Box Plot: Storage vs. Phone Model, created this plot to show the relationship between the two models and has a corelation

# Box plot for price per phone brand
ggplot(phone, aes(x = phone_brand, y = price)) +
  geom_boxplot() +
  labs(
    title = "Box Plot of Price per Phone Brand",
    x = "Phone Brand",
    y = "Price"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#4 Bar Plot: Price range vs. display size, made a stacked bar plot to show the high and the low price ranges within the graph
ggplot(phone, aes(x = price_range, y = Display_Size, fill = price_range)) +
  geom_boxplot() +
  labs(title = "Display Size Across Price Ranges",
       x = "Price Range",
       y = "Display Size (inches)") +
  theme_minimal() +
  theme(legend.position = "none")
```
## **6. Reproducibility**
